
This paper presented GPAC, a four-layer hierarchical controller for decentralized cooperative aerial transport on $\SEthree \times (\Sph^2)^N$. Each quadrotor runs an identical control and estimation stack using only local sensors and its cable attachment---no knowledge of $N$ or $m_L$ is required, and no payload, cable, or adaptive-parameter states are exchanged (only neighbor GPS positions for collision avoidance). Concurrent learning enables mass estimation without persistent excitation (converging in ${\sim}8$\,s, +33\% over gradient-only), and a modular CBF safety filter enforces operational constraints with ISSf guarantees provably compatible with the geometric attitude controller (Theorem~\ref{thm:compatibility}). High-fidelity Drake simulation with flexible bead-chain cables, ESKF sensor fusion, and Dryden wind validates 22.9\,cm tracking RMSE at under 1\,MFLOP/s per agent.

The hazard-oriented architecture---where each layer directly mitigates a specific transport risk (Table~\ref{tab:failure_modes})---combines formal stability certificates, adaptive learning, and runtime constraint enforcement without centralized coordination. The hierarchical timescale separation limits fault propagation, the decentralized implementation eliminates single points of failure, and the per-agent policy is invariant to team size by construction. These properties position GPAC toward assurance-oriented deployment of multi-UAV transport systems.

\textit{Limitations.} All results are simulation-only; hardware validation is the immediate next step. The shared broadcast trajectory is a single point of failure. The decentralized load estimator exhibits a $4\times$ gap versus the centralized baseline (49.5 vs.\ 12.4\,cm), motivating distributed consensus-based estimation. Agent dropout and dynamic reconfiguration are not addressed. The sequential CBF projection can be suboptimal when multiple constraints activate simultaneously.

\textit{Future work.} Flight experiments with 3--6 quadrotors, distributed estimation to narrow the centralized gap while preserving communication efficiency, agent dropout detection and $N$-adaptive mass re-estimation, and full QP-based multi-constraint resolution for the safety filter.
