% !TEX root = ../Main.tex
The GPAC framework is a four-layer hierarchical controller that assigns each physical degree of freedom to a control layer operating at a progressively higher bandwidth (Fig.~\ref{fig:architecture}). Each agent~$i$ receives only: (i)~a shared reference trajectory~$p_d(t)$ (broadcast before flight), (ii)~its own IMU/GPS/barometer measurements, and (iii)~local cable tension $T_i$ and direction $q_i$. The position of the load is obtained via GPS broadcasts at 10\,Hz. No other state of the payload, no other agent's cable state, and no centralized coordinator are required.

\begin{figure}[t!]
\centering
\resizebox{0.92\columnwidth}{!}{%
\begin{tikzpicture}[
  block/.style={draw, rounded corners=2pt, minimum width=3.0cm, minimum height=0.65cm, font=\footnotesize, align=center},
  arr/.style={-{Stealth[length=2mm]}, thick},
  every node/.style={font=\footnotesize}
]
  \node[block, fill=blue!10] (L1) {Layer 1: Position\\+ Anti-Swing (50\,Hz)};
  \node[block, fill=red!10, below=0.35cm of L1] (CBF) {CBF Safety Filter\\(200\,Hz)};
  \node[block, fill=blue!10, below=0.35cm of CBF] (L2) {Layer 2: Attitude\\on SO(3) (200\,Hz)};
  \node[block, fill=teal!10, right=1.2cm of L1] (L3) {Layer 3: CL\\Estimator (50\,Hz)};
  \node[block, fill=teal!10, right=1.2cm of CBF] (L4) {Layer 4:\\ESO ($\omega_0\!=\!50$)};
  \node[block, fill=green!8, right=1.2cm of L2] (ESKF) {ESKF Sensor\\Fusion (200\,Hz)};
  \node[block, fill=gray!10, below=0.35cm of L2, minimum width=7.2cm] (plant) {Quadrotor$_i$ + Bead-Chain Cable + Payload};
  \draw[arr] (L1) -- node[left, font=\scriptsize]{$F_{\text{des}},\; R_d$} (CBF);
  \draw[arr] (CBF) -- node[left, font=\scriptsize]{$f_{\text{safe}},\; R_d^{\text{safe}}$} (L2);
  \draw[arr] (L2) -- node[left, font=\scriptsize]{$\tau_i,\; f_i$} (plant);
  \draw[arr] (L3) -- node[above, font=\scriptsize]{$\hat{\theta}_i$} (L1);
  \draw[arr] (L4) -- node[above, font=\scriptsize]{$\hat{d}_i$} (CBF);
  \draw[arr] (ESKF) -- node[above, font=\scriptsize]{$\hat{p},\hat{v}$} (L4);
  \draw[arr, dashed] (ESKF.north) -- ++(0,0.55) -| (L1.east);
  \draw[arr, dashed, gray] (plant.east) -- ++(0.3,0) |- node[right, near start, font=\scriptsize, text=gray]{sensors} (ESKF.south);
\end{tikzpicture}
}
\caption{GPAC per-agent architecture. Blue: control cascade. Red: CBF safety overlay. Teal: estimation. Green: sensor fusion. Each agent executes this pipeline independently; no payload, cable, or estimator states are exchanged (only neighbor GPS positions for collision avoidance).}
\label{fig:architecture}
\end{figure}

\subsection{Layer 1: Position Tracking and Anti-Swing Control}

The outermost layer (${\sim}50$\,Hz) computes a desired thrust vector: $F_{\text{des},i} = F_{\text{fb},i} + F_{\text{ff},i} + F_{\text{cable},i} + F_{\text{swing},i} + F_{\text{eso},i}$, combining five physically motivated components. The PID trajectory tracking feedback (with per-axis anti-windup at $|e_{I,k}| \leq 2.0$) is $F_{\text{fb},i} = -K_p\,e_{p_i} - K_d\,e_{v_i} - K_i\,e_{I_i}$,with $K_p\!=\!\diag(6,6,8)$, $K_d\!=\!\diag(8,8,10)$, $K_i\!=\!\diag(0.1,0.1,0.2)$. The feedforward term compensates gravity and the nominal trajectory: $F_{\text{ff},i} = m_Q(a_{d_i} + ge_3)$.

Cable tension compensation prevents the position controller from treating cable force as a disturbance: $F_{\text{cable},i} = \kappa_T(t)\,T_i\,q_i$, where $\kappa_T(t) = \min\!\bigl(1,\,(t - t_{\text{taut},i})/\Delta_{\text{ramp}}\bigr)$,
and $t_{\text{taut},i}$ is the time cable~$i$ first reaches $T_i \geq 1.0$\,N and $\Delta_{\text{ramp}} = 2$\,s. The gain ramp gradually enables tension compensation using only the \emph{measured} cable force---no knowledge of $m_L$ or $N$ is required. This prevents impulsive loading during the slack-to-taut transition while remaining fully consistent with the decentralized estimation pipeline.

Anti-swing control suppresses pendular oscillations intrinsically on $\Sph^2$, i.e. $F_{\text{swing},i} = k_q\,e_{q_i} - k_\omega\,\omega_{q_i}$,
with $k_q\!=\!4.0$, $k_\omega\!=\!2.0$ ~\cite{lee2018geometric}. Here $e_{q_i}$~\eqref{eq:eq} steers $q_i \to q_{d_i}$ along the geodesic and $\omega_{q_i} \in T_{q_i}\Sph^2$ is the cable angular velocity. Together, these terms reduce $\Psi_{q_i}$ monotonically in the absence of external perturbation. The ESO disturbance estimate (Layer~4) is fed forward as $F_{\text{eso},i} = m_Q\hat{d}_i$.

The desired rotation $R_{d_i} \in \SOthree$ is extracted from $F_{\text{des},i}$ by aligning the body $z$-axis with the thrust direction~\cite{lee2010geometric}:
\begin{equation}
  b_{3c} = \frac{F_{\text{des},i}}{\norm{F_{\text{des},i}}},\;\; b_{2c} = \frac{b_{3c} \times b_{1d}}{\norm{b_{3c} \times b_{1d}}},\;\; b_{1c} = b_{2c} \times b_{3c},
  \label{eq:Rd_extract}
\end{equation}
with $b_{1d} = [\cos\psi_d,\, \sin\psi_d,\, 0]^\top$ setting the desired heading.

\subsection{Layer 2: Geometric Attitude Control on SO(3)}

The inner attitude loop (200\,Hz) applies the geometric tracking controller~\cite{lee2010geometric}:
\begin{equation}
  \tau_i = -k_R\,e_{R_i} - k_\Omega\,e_{\Omega_i} + \Omega_i \times J\Omega_i,
  \label{eq:att_ctrl}
\end{equation}
with $k_R\!=\!8.0$, $k_\Omega\!=\!1.5$, and angular velocity error $e_{\Omega_i} = \Omega_i - R_i^\top R_{d_i}\Omega_{d_i}$. The first two terms provide proportional-derivative control on $\SOthree$; the third compensates gyroscopic coupling. The feedforward terms $\hatmap{\Omega_i}R_i^\top R_{d_i}\Omega_{d_i} - R_i^\top R_{d_i}\dot{\Omega}_{d_i}$ are omitted under the simplification $\Omega_{d_i} \approx 0$ (valid when the desired thrust direction varies slowly relative to the 200\,Hz attitude bandwidth).

\begin{proposition}[Almost-global exponential stability~\cite{lee2010geometric}]\label{prop:lyap}
Define the Lyapunov function
\begin{equation}
  V_R = \tfrac{1}{2}k_R\,\Psi_{R_i} + \tfrac{1}{2}e_{\Omega_i}^\top J\,e_{\Omega_i} + c\,e_{R_i}^\top J\,e_{\Omega_i},
  \label{eq:VR}
\end{equation}
where $\Psi_{R_i} = \frac{1}{2}\tr(I - R_{d_i}^\top R_i)$ is the attitude configuration error and $c>0$ is sufficiently small. Under~\eqref{eq:att_ctrl} with $\Psi_{R_i}(0)<2$:
\begin{equation}
  \dot{V}_R \leq -\lambda_{\min}(W)\bigl(\norm{e_{R_i}}^2 + \norm{e_{\Omega_i}}^2\bigr),
  \label{eq:VR_dot}
\end{equation}
for a positive-definite $W = W(k_R, k_\Omega, J)$, yielding exponential convergence on the dense open set $\{\Psi_R < 2\} \subset \SOthree$.
\end{proposition}

\subsection{Layers 3--4: ESO and Concurrent Learning Estimator}

\textit{Layer~4} is a third-order Extended State Observer (ESO) per translational axis, modeling the dynamics as a double integrator with lumped disturbance $\ddot{p}_k = b_0 u_k + d_k(t)$, $b_0 = 1/m_Q$:
\begin{equation}
  \dot{\hat{z}}_{1k} = \hat{z}_{2k} + 3\omega_0\tilde{x}_k,\;\; \dot{\hat{z}}_{2k} = \hat{z}_{3k} + 3\omega_0^2\tilde{x}_k + b_0 u_k,\;\; \dot{\hat{z}}_{3k} = \omega_0^3\tilde{x}_k,
  \label{eq:eso}
\end{equation}
with $\tilde{x}_k = x_k - \hat{z}_{1k}$ and all poles at $-\omega_0 = -50$\,rad/s (settling time ${\sim}0.1$\,s). The disturbance estimate $\hat{d}_i = [\hat{z}_{3x}, \hat{z}_{3y}, \hat{z}_{3z}]^\top$ is saturated to $|\hat{d}_{i,k}| \leq 20$\,m/s$^2$ and fed forward via $F_{\text{eso},i}$. Under bounded $\dot{d}$, the estimation error scales as $O(\sup|\dot{d}|/\omega_0)$~\cite{guo2013active}.

\textit{Layer~3} adaptively estimates $\hat{\theta}_i \approx m_L/N$ via concurrent learning~\cite{chowdhary2010concurrent}, providing the gravity compensation term $\hat{\theta}_i(ge_3 + \ddot{p}_d^L)$ to Layer~1. The full derivation is given in Section~\ref{sec:estimation}.

\subsection{Cascade Stability and Fault Containment}

The four-layer cascade exploits deliberate timescale separation: attitude ($k_R/J \approx 200$\,rad/s) $\gg$ ESO ($\omega_0 = 50$\,rad/s) $\gg$ position ($K_p/m_Q \approx 5$\,Hz). Standard singular perturbation arguments~\cite{khalil2002nonlinear} guarantee composite stability when this bandwidth ratio exceeds a computable threshold. The hierarchical structure also provides fault containment: mass estimation errors (Layer~3) affect only the feedforward in Layer~1 without corrupting attitude tracking; ESO transients are filtered before reaching the attitude loop; and CBF interventions are bounded by the tilt constraint (Section~\ref{sec:safety}). This modularity supports independent verification of each layer~\cite{koopman2017autonomous}.
